Criticisms of the IHME COVID-19 model

### 2020-04-24 text/20/ihme-covid-model.Rmd
Continued work on ihme-covid-model.

There is a wonderful sense of openness and transparency about research these days, and the IHME model is a good example of this.

The Institute for Health Metrics and Evaluation has been in business since 2007, so they are no Johnny-Come-Lately operation. The page that lists their staff members has well over 400 pictures. Not all of these people are working on COVID-19, I'm sure, but it appears to be a solid team effort.

The IHME is documented in several places.

@@

Murray CJL. Forecasting COVID-19 impact on hospital bed-days, ICU-days, ventilator-days and deaths by US state in the next 4 months. MedRxiv preprint, 2020-04-22. [pdf format](https://www.medrxiv.org/content/10.1101/2020.03.27.20043752v1.full.pdf).

@@

@@

>>>>

To avoid too much confusion, I will use the term "IHME model" to refer to the COVID-19 prediction model developed by the Institute for Health Metrics and Evaluation and reserve the term "the Institute" to refer to the organization itself.

The Institute has been in business since 2007, so they are no Johnny-Come-Lately operation. The page that lists their staff members has well over 400 pictures. Not all of these people are working on COVID-19, I'm sure, but it appears to be a solid team effort.

There is a wonderful sense of openness and transparency about research these days, and the Institute is a good example of this. The Institute shares the code and (when it can) the data on their [github site](https://github.com/ihmeuw). This allows other researchers to borrow their work and easily adapt it as needed. When the Institute makes changes to their work, anyone can quickly download these updates.

@@

Murray CJL. Forecasting the impact of the first wave of the COVID-19 pandemic on hospital demand and deaths for the USA and European Economic Area countries. Preprint, 2020-04-21. Available in [pdf format](http://www.healthdata.org/sites/default/files/files/Projects/COVID/RA_COVID-forecasting-USA-EEA_042120.pdf).

+ At the time of the writing of this webpage, this was the most current description of the IHME model.

Murray CJL. Forecasting COVID-19 impact on hospital bed-days, ICU-days, ventilator-days and deaths by US state in the next 4 months. Preprint, 2020-03-26. Available in [pdf format](https://www.medrxiv.org/content/10.1101/2020.03.27.20043752v1.full.pdf).

+ This article describes an earlier model.

Xu B, Gutierrez B, Mekaru S, Sewalk K, Goodwin L, Loskill A, Cohn EL, Hswen Y, Hill SC, Cobo MM, Zarebski AE, Li S, Wu C-H, Hulland E, Morgan JD, Wang L, O'Brien K, Scarpino SV, Brownstein JS, Pybus OG, Pigott DM, Kraemer MUG. Epidemiological data from the COVID-19 outbreak, real-time case information. Scientific Data, 2020-03-24. doi:10.1038/s41597-020-0448-0. Available in [html format](https://www.nature.com/articles/s41597-020-0448-0) or [pdf format](https://www.nature.com/articles/s41597-020-0448-0.pdf).

+ This article describes how the initial datasets used by the IHME model were collected. In a rapidly changing environment, some of the details may have changed, and at least one of the links in the article points to a gitrhub page that no longer exists. But most of the links are good. The breadth of the data collection is very impressive and the authors do a good job documenting everything.

@@

+ This webpage talks in non-technical terms about some of the choices that the researchers chose for the IHME model.

@@

+ The focus of this article is on the White House press briefing and the uncertain source of some of the numbers used in that briefing. There is, however, a good amount of detail about the IHME model.
